---
title: "Activity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load and process data
```{r}
data <- read.csv("activity.csv")
data <- data[ with (data, { !(is.na(steps)) } ), ]
```
## calculate mean total numbers of steps taken per day
```{r}
library(dplyr)
data1 <- group_by(data, date)
stepbyday <- summarise(data1, total =sum(steps))
hist(stepbyday$total, main = "Histogram of total number of steps per day")
```
## mean and median steps taken by day
```{r}
summary(stepbyday)
```
## time series plot of the average number of steps taken
```{r}
stepbyinterval <- aggregate(steps ~ interval, data, mean)
plot(stepbyinterval$interval, stepbyinterval$steps, type = 'l', 
     xlab = 'Interval', ylab = 'average total steps', 
     main = "average numbers of steps over days")
```
## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
row <- which.max(stepbyinterval$steps)
stepbyinterval[row, ]
```
## Code to describe and show a strategy for imputing missing data
```{r}
sum(is.na(data))
newdata <- data
for (i in 1:nrow(newdata)) {
  if (is.na(newdata$steps[i])) {
    intervalX <- newdata$interval[i]
    stepsX <- stepbyinterval[stepbyinterval$interval == intervalX, ]
    newdata$step[i] < stepsX$steps
  }
}
```
## Histogram of the total number of steps taken each day after missing values are imputed
```{r}
stepbyday2 <- aggregate(steps ~ date, newdata, sum)
hist(stepbyday2$steps,  main = "total steps taken per day after missing values are imputed")
mean(stepbyday2$steps)
median(stepbyday2$steps)
mean(stepbyday$total)
median(stepbyday$total)
```
## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
newdata$Day <- weekdays(as.Date(newdata$date))
newdata$Daytype <- "weekday"
newdata$Daytype[newdata$Day %in% c("Saturday", "Sunday")] <- "weekend"
                        
stepbyinterval_new <- aggregate(steps ~ interval + Daytype, newdata, mean)
library(ggplot2)
qplot(interval, steps, data = stepbyinterval_new,
      type = 'l', geom = c("line"), 
      xlab = 'interval', ylab = 'Average number of steps',
      main = "") + 
      facet_wrap(~ Daytype, ncol = 1)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
